---
title: "Changing fonts and enhancing plots"
author: "Ashwin Malshe"
date: "1 October 2024"
output: html_document
---

For using any fonts outside the fonts installed on your system, I strongly recommend using `showtext` package.

Install is by running this code in the console:

install.packages("showtext")

```{r}
#| label: setup
#| include: false

library(ggplot2)
library(showtext)
```


Get the font you like either from Google Fonts or from a `.ttf` or `.otf` file.

I am importing this cool font:
https://fonts.google.com/specimen/Monoton

The `family` argument below is a name that I choose to give to this font to be used in this session. You can use any name you want.

```{r}
sysfonts::font_add_google(name = "Monoton", family = "Monoton", db_cache = FALSE)
```


Before we create a plot, we need to start `showtext`:

```{r}
showtext_auto()
```

Now we are ready to create a plot

```{r}
#| message = FALSE
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "A plot using Monoton") +
  theme(plot.title = element_text(family = "Monoton", color = "blue", hjust = 0.5))

```


## Reading a local font file

```{r}
sysfonts::font_add(family = "retina", regular = "Retina-Light.otf")
```


```{r}
#| message = FALSE
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(title = "A plot using Monoton", 
       subtitle = "With a subtitle using Retina Light",
       x = "Engine Displacement (liters)",
       y = "Highway Mileage (miles/gallon)") +
  theme(plot.title = element_text(family = "Monoton"),
        plot.subtitle = element_text(family = "retina"),
        axis.title = element_text(family = "retina"),
        axis.text = element_text(family = "retina"))

```
## Enhancing the plot


Let's annotate a point on extreme right.

```{r}
mpg |> 
  select(displ, hwy, model) |> 
  filter(displ == 7)
```





```{r}
#| message = FALSE
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  annotate(geom = "label", x = 6.8, y = 27, label = "Corvette", family = "retina") +
  # annotate(geom = "point", x = 7, y = 24, shape = 21, size = 3) +
  # annotate(geom = "point", x = 7, y = 24, color = "red") +
  labs(title = "A plot using Monoton", 
       subtitle = "With a subtitle using Retina Light",
       x = "Engine Displacement (liters)",
       y = "Highway Mileage (miles/gallon)") +
  theme(plot.title = element_text(family = "Monoton"),
        plot.subtitle = element_text(family = "retina"),
        axis.title = element_text(family = "retina"),
        axis.text = element_text(family = "retina"))

```

How about adding a segment connecting the point and the label? We can use `geom_segment` or `annotate` to achieve this.

```{r}
#| message = FALSE
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  annotate(geom = "label", x = 6.6, y = 30, label = "Corvette", family = "retina") +
  annotate(geom = "segment", x = 6.7, y = 29, xend = 7, yend = 24, linewidth = 0.2) +
  annotate(geom = "point", x = 7, y = 24, shape = 21, size = 3) +
  annotate(geom = "point", x = 7, y = 24, color = "red") +
  labs(title = "A plot using Monoton", 
       subtitle = "With a subtitle using Retina Light",
       x = "Engine Displacement (liters)",
       y = "Highway Mileage (miles/gallon)") +
  theme(plot.title = element_text(family = "Monoton"),
        plot.subtitle = element_text(family = "retina"),
        axis.title = element_text(family = "retina"),
        axis.text = element_text(family = "retina"))

```

