---
title: "ggplot2 Basics"
author: "Ashwin Malshe"
date: "23 September 2025"
output: html_document
editor: source
execute: 
  echo: true
---

::: callout-note
In the YAML header above, take a note of the following options that we did not see before -

```yaml
---
editor: source
execute: 
  echo: true
---
```
The first option will set your editor to`source`, which is the standard coding editor where we combine markdown and R. If you change that to `visual` it will show the visual editor. You can manually change this setting by toggling the buttons on top of YAML header. This does not change the behavior of your code.

The second argument `execute` is important and it will change the way the rendering will work. Currently, I have specified `echo: true` which means the R code and its output will be visible to you. However, if you are creating a report for people who may not be interested in the code, you can hide it using `echo: false`. **This affects only the rendered output, be it html or pdf.** By the way, this is how I created your homework assignment.

There are more options available within `execute` and you should check them out here: https://quarto.org/docs/reference/formats/html.html#execution

Also, learn about markdown here: https://quarto.org/docs/authoring/markdown-basics.html
:::

```{r}
#| label: setup
#| include: false

library(tidyverse)
```

## `mpg` dataset

Let's start with a dataset named `mpg` that comes bundled with `ggplot2`

Get the variable names (i.e., column names)

```{r}
names(mpg)
```

Check the `class` of each variable in `mpg`

```{r}
sapply(mpg, class)
```

Take a glimpse at the variables in `mpg`

```{r}
glimpse(mpg)
```

## The structure of a `ggplot2` plot

`ggplot2` library uses a "layered" grammar of graphics. A plot is created layer by layer. The base layer is the canvas. Next `ggplot2` lays down axes on top of the canvas. Next, it lays down geometric shapes (called as `geoms`), and so on. **The ordering of the `geoms` matters.**

```{r}
#| eval: false 

ggplot(data = ,
       mapping = aes()) +
  geom_*() +
  scale_*() +
  labs() + 
  facet_*() +
  theme()

```

## `ggplot()` function

`ggplot()` is the key function in `ggplot2`. If you execute the following chunk, you will get a gray box as an output. This is a standard output from `ggplot()` in the absence of any information on the data.

```{r}
ggplot()
```

`ggplot()` uses `data` and `mapping` arguments to create the base layer for a plot. Giving `data` alone won't output anything informative.

```{r}
ggplot(data = mpg)
```

We need to provide the mapping from the data to aesthetics. This is not always required because some `geoms` will have their own default statistics mapping to the aesthetic. Consider this as a case where a `geom` creates a variable from the data and then uses that to make the plot.

In the next chunk, we provide the mapping from `cty` to the X axis and `hwy` to the Y axis. This will lead to a more informative plot compared to just a large gray box we created above!

```{r}
ggplot(data = mpg,
       mapping = aes(x = cty, y = hwy))
```

We still don't know what we want to show on the plot. However, now we can start to think about the possible visualizations. Both `cty` and `hwy` are continuous variables. So making a scatterplot to show their relationship will be informative.

We do that by **adding** `geom_point()` to `ggplot()`

Note two important aspects here:

1.  We add a layer to a `ggplot` by using a `+` sign. This is not a `|>`! People make this mistake very often, so be careful here.

2.  We don't provide any `data` or `mapping` to `geom_point()`. This is because all the layers inherit these arguments from `ggplot()`. So you don't have to keep on specifying them in each layer.

*Important: You can specify `data` and `mapping` inside any `geom_`. It will not be passed on to any other layer. It will remain confined to that layer alone.*

## Inheritance explained

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point()
```

This is the same as above:

```{r}
ggplot() +
  geom_point(data = mpg, mapping = aes(x = cty, y = hwy))
```

Note that we can add another layer and it will inherit the data and mapping from `ggplot` again.

```{r}
mpg |> 
ggplot(mapping = aes(x = cty, y = hwy)) +
  geom_point() +
  geom_smooth()
```

But if `ggplot()` doesn't specify them, subsequent layers don't inherit anything from another layers. So in the code below, the line doesn't show.

```{r}
ggplot() +
  geom_point(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_smooth()
```

Thus, the arguments passed to `ggplot()` are **global** while the ones passed to a `geom_*()` are **local**.
