---
title: "Dataframe manipulations"
format: html
editor: source
---

Install `tidyverse` package by typing the following command in the console. **Do not execute it inside this file!**

install.packages("tidyverse")

```{r}
library(tidyverse)
```

Import outside data into R

```{r}
d1 = read_csv("company_financials_csv.zip")
```

Take a look at the "variables_description.xlsx" file to understand the variables.

### dataframe vs tibble

Take a look at the class of `d1`

```{r}
class(d1)
```

The class contains multiple subclasses. More importantly, it is a dataframe as evident from the `data.frame` value. `tidyverse` packages convert R dataframe into a customized format called `tibble`. It is just a dataframe with a few additional attributes. It is not crucial to know the minor differences between a tibble and a dataframe at this point and we will treat tibble as any other dataframe.


## Explore the dataframe

`glimpse()` function gives us a glimpse of the dataframe


```{r}
glimpse(d1)
```

`head()` function allows you to print first few rows of the dataframe. The default is 6 rows but you can put any number you want inside the function

```{r}
head(d1)
```

```{r}

# prints first 10 rows

head(d1, 10)
```



`summary()` function will provide basic summaries of all numeric variables

```{r}
summary(d1)
```

### Subsetting a dataframe

Similar to a vector, you can use `[]` to extract specific elements from a datframe. For example, if you want to extract the value at the intersection of the 3rd row and 12th column, use this:

```{r}
d1[3, 12]
```

Notice that the resulting object is not just a single value but looks like a dataframe. Let's find out its class.

```{r}
class(d1[3, 12])
```

So this is a tibble. What id we wanted just the text string instead of a tibble? Here is how you can subset it:

```{r}
d1[[12]][3]
```

Or if you know the name of the 12th column,

```{r}
d1$conm[3]
```

Both `[[]]` and `$` operators return the column as a vector. Next, we use `[]` to subset the vector and return a string.


```{r}
class(d1$conm[3])
```

## pipes

Pipes are one of th emost critical aspects of R. This smoothes data wrangling and allows you to write code that is easily readable.

This is a pipe: |>

**The keyboard shortcut for pipe is "Shift + Command + M" on Macs and "Shift + Control + M" on Windows.**

A pipe takes the object on its left and assigns it to the first argument on the function on right

```{r}
sum(2, 5)

# can be written as this using a pipe

2 |> sum(5)

```

Similarly,

```{r}
class(d1)

# can be written as 

d1 |> class()
```


## Typical operations for dataframes

Here are the most common operations we do on a dataframe:

1. Select columns (variables)
2. Select rows (observations)
3. Rename variables
4. Create new variables
5. Create summary variables
6. Join dataframes
7. Change the structure of a dataframe from wide to long or long to wide

We will predominatly use `dplyr` package, which is a part of `tidyverse` to achieve this task.

## Select columns

`select()` function from `dplyr` is the most powerful function for this task.


The following function will return a dataframe with first five columns of `d1`. 

`select(d1, 1:5)`

But if you execute it, it will print the entire dataframe. Instead, let's print only first 6 rows.


```{r}
head(select(d1, 1:5))
```

In this case, we used a nested function call. `select()` is nested inside `head()`. In this case, the function inside the parantheses gets executed first and then the outer function.

We can instead use a pipe:

```{r}
select(d1, 1:5) |> 
  head()
```


Or even two pipes

```{r}
d1 |> 
  select(1:5) |> 
  head()
```


