---
title: "Homework 2"
author: "Justin Gould"
date: "October 17, 2025"
format:
  html:
    theme: zephyr
    embed-resources: true
execute:
  eval: true
  echo: true
  message: false
---

```{r}
#| warning: false 
#| message: false
#| echo: true

pacman::p_load(tidyverse, ggtext, png, jpeg)
theme_set(theme_minimal())
setwd("C:/Users/justi/OneDrive/Desktop/Grad School/UTSA 1st semester/DA 6233 - Data Analytics Visualization and Communication/Homework 2")

```

Homework 2 focuses on developing your `ggplot2` skills. You will work with the `company_financials_csv.zip` dataset that you explored in Homework 1. Your familiarity with the data structure and variable meanings from the previous assignment will be valuable as you create various visualizations.

```{r}
#| warning: false 
#| message: false
#| echo: true

d1 = read_csv("tnhc7nsnznjqy9gp.csv", show_col_types = FALSE) |>
  filter(saleq > 0 & atq > 0) |>
  mutate(conm = stringr::str_to_title(conm), # Converts the string to title case
         datadate = lubridate::ymd(datadate)) # Convert datadate into a date variable
```

<p style="border:3px; border-style:solid; border-color:#b50000; padding: 1em;">

This homework consists of 7 questions. <b>Your objective is to reproduce the output shown in the HTML file for Q1 through Q7.</b>

</p>

{{< pagebreak >}}

## Q1 (2 points)

Create a bar graph showing the **total assets** (`atq`) for each company, arranged in descending order. I'm providing the code to prepare the dataset:

```{r}
#| warning: false 
#| message: false
#| echo: true

d1_1 = d1 |>
  group_by(conm) |>
  summarize(total_assets = sum(atq), .groups = "drop")
```

Use `d1_1` to create the bar plot. Make sure the bars are ordered and the axes are properly labeled.

[Pay attention to the ordering of bars, axis formatting with dollar signs, and descriptive axis titles.]{.aside}

```{r}
#| warning: false 
#| message: false
#| echo: true

d1_1 |> 
    ggplot(aes(x = total_assets, y = reorder(conm, -total_assets))) +
    geom_col() +
    labs(x = "Total Assets in $ Millions", y = "Company") + scale_x_continuous(labels = scales::label_currency())

```

{{< pagebreak >}}

## Q2 (1 point)

Enhance the plot from Q1 by adding text labels inside the bars showing the total assets in billions (rounded to 1 decimal place). Use `hjust = 1.1`, `size = 3`, and `fontface = "bold"` for the text labels. The exact positioning may vary slightly from the reference figure.

```{r}
#| warning: false 
#| message: false
#| echo: true

d1_1 |> 
    ggplot(aes(x = total_assets, y = reorder(conm, -total_assets))) +
    geom_col() +
    geom_text(aes(label = round(total_assets / 1000000, 1)), color = "white", hjust = 1.1, size = 3, fontface = "bold") +
    labs(x = "Total Assets in $ Millions", y = "Company") + scale_x_continuous(labels = scales::label_currency())
```

{{< pagebreak >}}

## Q3 (2 points)

Let's explore the relationship between company size (measured by total assets) and profitability (measured by return on assets). Create two variables using `mutate`:

`roa = oiadpq / atq`

`log_assets = log(atq)`

```{r}
#| warning: false 
#| message: false
#| echo: true
d1_2 = d1 |> mutate(roa = oiadpq / atq, log_assets = log(atq))|>
  filter(!is.na(roa), !is.na(log_assets))
```

Then create a scatter plot with `log_assets` on the X axis and `roa` on the Y axis. Filter out any rows where `roa` is `NA` to avoid warnings.

[Do larger companies generate better returns? The relationship might surprise you!]{.aside}

```{r}
#| warning: false 
#| message: false
#| echo: true
d1_2 |> 
    ggplot(aes(x=log_assets, y=roa)) +
    geom_point(shape = 21, color = "black", fill = "steelblue", size = 2, alpha = 0.7) +
    geom_smooth(method = "loess", color = "orangered") +
    labs(x = "Log of Total Assets", y = "Return on Assets")
```

{{< pagebreak >}}

## Q4 (2 points)

Compare the **asset turnover ratio** (sales divided by total assets) over time for six tech giants: Apple, Google (Alphabet), Microsoft, Amazon, Meta, and Tesla. Use `fyearq` on the X axis.

Here's the code to prepare the data:

```{r}
#| warning: false 
#| message: false
#| echo: true
d1_4 = d1 |>
  filter(tic %in% c("AAPL", "GOOGL", "MSFT", "AMZN", "META", "TSLA")) |>
  mutate(asset_turnover = saleq / atq,
         fyearq_qtr = lubridate::yq(paste(fyearq, fqtr, sep = "-")))
```

Now use `d1_4` to create the following plot.

[Asset turnover measures how efficiently a company uses its assets to generate sales.]{.aside}

```{r}
#| message: false
#| echo: true
#| warning: = false


d1_4 |> ggplot(aes(fyearq, asset_turnover)) +
    geom_line(color = "navyblue", size = 1.10) +
    labs(x = "Date", y = "Asset Turnover Ratio") +
    facet_wrap(~ factor(tic, levels = c( "GOOGL", "AMZN", "AAPL", "META", "MSFT", "TSLA")), 
               ncol = 2, 
               labeller = as_labeller(c(
                   AAPL = "Apple Inc",
                   AMZN = "Amazon.com Inc",
                   META = "Meta Platforms Inc",
                   MSFT = "Micrsoft Corp",
                   TSLA = "Tesla Inc",
                   GOOGL = "Alphabet Inc")))
```

{{< pagebreak >}}

## Q5 (3 points)

Nvidia has experienced tremendous growth in both revenue and market capitalization. Let's visualize these two metrics together over time.

First, create a properly formatted dataset. The last line in the code shows that the values are converted to log10. Here's the code:

```{r}
#| warning: false 
#| message: false
#| echo: true
d1_5 = d1 |>
  filter(conm == "Nvidia Corp") |>
  mutate(mkt_cap = prccq * cshoq) |> # Create market capitalization
  filter(mkt_cap > 0) |>
  select(conm, datadate, mkt_cap, saleq) |>
  pivot_longer(cols = c(mkt_cap, saleq),
               names_to = "metric",
               values_to = "value") |> 
  mutate(value = log10(value))
```

Examine `d1_5` in the console using `head()` to understand its structure. Then create a line plot with `datadate` on the X axis.

I have used the following colors in the plot: #FF0066, #FF9A00, and #9112BC.

[The Y axis plots log values. However, it is difficult for people to understand what they mean. Therefore, I am using sensible labels as you see in the plot. You will have to use log10 of these labels as your Y axis breaks. If you have difficulties, write to me.]{.aside}

```{r}
#| warning: false 
#| message: false
#| echo: true

d1_5 |> 
  ggplot(aes(x = datadate, y = value)) +
  geom_line(aes(color = metric), size = 1.5, na.rm = TRUE) +
  scale_color_manual(
    name = "Financial Metric",
    values = c("mkt_cap" = "#FF0066", "saleq" = "#FF9A00"),
    labels = c("mkt_cap" = "Market Cap", "saleq" = "Sales")
  ) +
  scale_y_continuous(
    breaks = c(2, 3, 4, 5, 6, log10(3000000)),
    labels = function(x) scales::label_currency()(10^x)
  ) +
  labs(
    title = "Nvidia's <span style='color:#FF0066;'>Market Cap</span> and <span style='color:#FF9A00;'>Sales</span> both surge",
    subtitle = "All amounts in <span style='color:#9112BC;'>million USD</span>",
    x = "Date",
    y = NULL,
    color = "Financial Metric"
  ) +
  scale_x_date(date_labels = "%Y") +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "top",
    legend.title = element_text(),
    legend.text = element_markdown(),
    plot.title = element_markdown(size = 16),
    plot.subtitle = element_markdown(size = 12)
  )

```

{{< pagebreak >}}

## Q6 (3 points)

Create a side-by-side bar plot comparing **total annual R&D spending** (`xrdq`) for Apple and Microsoft starting from 2010. Use the Lobster font from Google Fonts throughout the plot. Set the bar colors manually to: `c("#ff6b35", "#004e89")`.

[Hint: Use `position_dodge()` inside `geom_col()` to create side-by-side bars.]{.aside}

```{r}
library(showtext)
library(sysfonts)
font_add_google("Lobster", "lobster")

```

```{r}
#| warning: false 
#| message: false
#| echo: true


showtext::showtext_auto()

d1 |> 
    filter(tic %in% c("AAPL", "MSFT")) |> 
    mutate(xrdq_scaled = xrdq * (30000 / max(xrdq, na.rm = TRUE))) |>
    ggplot(aes(x = fyearq, y = xrdq_scaled, fill = tic)) + 
    geom_col(position = position_dodge(width = 0.8), width = 0.9) +
    scale_fill_manual(
        values = c("AAPL" = "#ff6b35", "MSFT" = "#004e89"),
        labels = c("AAPL" = "Apple", "MSFT" = "Microsoft")
    ) +
    scale_x_continuous(
        limits = c(2009.5, 2024.5),
        breaks = seq(2010, 2024, by = 1),
        minor_breaks = NULL
    ) +
    scale_y_continuous(
        limits = c(0, 30000), 
        breaks = c(10000, 20000, 30000),
        minor_breaks = seq(0, 30000, by = 5000),
        labels = c("10,000", "20,000", "30,000")
    ) +
    labs(
        fill = "",
        y = "R&D Spending ($ million)",
        x = "Fiscal Year"
    ) +
    theme_minimal(base_size = 14, base_family = "lobster") +
    theme(
        legend.position = "top",
        panel.grid.major.x = element_line(color = "gray90", linewidth = 0.3),
        panel.grid.major.y = element_line(color = "gray90", linewidth = 0.6),
        panel.grid.minor.y = element_line(color = "gray90", linewidth = 0.3)
    )

showtext::showtext_end()

```

{{< pagebreak >}}

## Q7 (2 points)

Create a bar graph showing total revenue (`saleq`) for all 13 companies in fiscal year 2022. Use Montserrat font from Google Fonts for the title background and Noto Sans for axis text. The title background color is #1E93AB, the plot background color is #DCDCDC, and bars are filled with #67C090. Add company logos for the top 3 revenue generators using `annotation_raster()`.

```{r}

font_add_google("Montserrat", "Noto Sans")
apple_img <- png::readPNG("official-apple-logo-png.png")
alphabet_img <- png::readPNG("alphabet-logo-rotated.png")
amazon_img <- png::readPNG("Amazon_icon.png")
```

```{r}
#| warning: false


showtext::showtext_auto()

d1_6 <- d1 |> 
    filter(fyearq == 2022) |>       
    group_by(conm) |>                
    summarize(total_sale = sum(saleq, na.rm = TRUE)) 

d1_6 |> 
    ggplot(aes(x = reorder(conm, -total_sale), y = total_sale)) + 
    geom_col(fill = "#67C090") +
    labs(
        title = "Amazon leads tech giants in total revenue in 2022",
        x = NULL,
        y = NULL
    ) +
    scale_y_continuous(
        limits = c(0, max(500000, 1.1 * max(d1_6$total_sale))), 
        breaks = c(0, 100000, 200000, 300000, 400000, 500000),
        minor_breaks = seq(0, 500000, by = 100000),
        labels = c("0", "100,000", "200,000", "300,000", "400,000", "500,000")
    ) +
    annotation_raster(
        amazon_img,    
        xmin = 1 - 0.45,
        xmax = 1 + 0.45,   
        ymin = d1_6$total_sale[1] * 1.45,
        ymax = d1_6$total_sale[1] * 1.75) +
    annotation_raster(
        alphabet_img,    
        xmin = 3 - 0.4,
        xmax = 3 + 0.4,   
        ymin = d1_6$total_sale[1] * .0,
        ymax = d1_6$total_sale[1] * 1.) + 
    annotation_raster(
        apple_img,
        xmin = 2 - 0.25,
        xmax = 2 + 0.25,   
        ymin = d1_6$total_sale[1] * 1.15,
        ymax = d1_6$total_sale[1] * 1.35 
    ) +
    theme_minimal(base_size = 14, base_family = "NotoSans") +
    theme(
        axis.text.x = element_text(angle = 30, hjust = .7),
        axis.text.y = element_text(),        
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.background = element_rect(fill = "#DCDCDC"),
        plot.title = element_textbox_simple(
            size = 16,
            padding = margin(5, 5, 5, 5),
            fill = "#1E93AB",
            color = "white",
            family = "Montserrat" 
        )
    )
```
