---
title: "Homework 3"
subtitle: "DA 6233"
author: "Justin Gould qbu192"
date: "October 31, 2025"
format: 
  html: 
    embed-resources: true
    theme: zephyr
execute: 
  echo: true
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(plotly)
library(highcharter)
library(lubridate)
library(scales)
```

```{r}
#| label: load-data
#| include: false

# I set the working directory first
setwd("C:/Users/justi/OneDrive/Desktop/Grad School/UTSA 1st semester/DA 6233 - Data Analytics Visualization and Communication/Homework 3/Homework 3/company_financials_csv")
#next, I read in the CSV
data = read.csv("tnhc7nsnznjqy9gp.csv")
#Next, I created a new dataframe called data2.0 with my new columns
data2.0 = data |>
  mutate('ROA (Return on Assets)' = oiadpq / atq,
         'Asset Turnover' = saleq / atq, 
         'Profit Margin' = round(oiadpq / saleq, 3),
         'year' = substring(datadate, 1,4),
         'Asset Growth' = (100 * (atq - lag(atq)) / lag(atq)))
```

## Variables in this Assignment

**ROA (Return on Assets)**: `oiadpq / atq` - Measures how efficiently a company uses its assets
**Asset Turnover**: `saleq / atq` - Measures how efficiently a company uses assets to generate sales
**Profit Margin**: `oiadpq / saleq` - Operating income *after* depreciation as percentage of sales

## Q1 (3 points)

Create an interactive scatter plot showing the relationship between Return on Assets (ROA) and Asset Growth for all companies, with animation by year.

1. Filter the data to include only years 2015-2023 and remove any rows with missing sales or asset growth values
2. Create a `ggplot` scatter plot with:
   - x-axis: log of `saleq` (Sales)
   - y-axis: `asset_growth` (Asset Growth %)
   - Color: `conm` (Company Name)
3. Add `geom_point()` with alpha = 0.7 and size = 3
4. Apply `theme_minimal()` and remove the legend
5. Convert to interactive plot using `ggplotly()`

Don't forget to include `frame`!

```{r}
#| label: Q1
#| warning: false
#| fig-width: 10
#| fig-height: 7

chart <- data2.0 |> 
    filter(year >=  2015, year <= 2023, !is.na(saleq),!is.na("Asset Growth")) |> 
    ggplot(aes(
        x =log(saleq), y=`Asset Growth`, color = conm)) + 
    geom_point(aes(size = 3, alpha = 0.7, frame = year, ids = saleq)) +
    scale_y_continuous(
        breaks = c(-50, 0, 50), limits = c(-50, 75)) +
    labs(
        title = "Tech Company Performance: Sale vs Asset Growth",
        x = "Log(Sales)", 
        y = "Asset Growth (%)") +
    theme_minimal() + 
    theme(
        legend.position = "none",
        panel.grid.major.x = element_line(color = "gray90", linewidth = 0.3),
        panel.grid.major.y = element_line(color = "gray90", linewidth = 0.6),
        panel.grid.minor.y = element_line(color = "gray90", linewidth = 0.3)
    )

ggplotly(chart)
```

---

## Q2 (3 points)

Create a heatmap showing average annual profit margin by company and year


1. Filter data to include years 2015-2023
2. Group by company (`conm`) and year, then calculate average profit margin
3. Round the result to 3 decimal place
4. Apply `hc_theme_538()` theme
5. Customize colors using `hc_colorAxis()` with a blue-to-orange color scheme. I used the following colors:
"#0d47a1", "#1976d2", "#42a5f5", "#ffb74d", "#ff8a65", "#ff5722"


```{r}
#| label: Q2
#| fig-width: 10
#| fig-height: 8

matrix <- data2.0 |> 
    filter(year >=  2015,
           year <= 2023) |> 
    group_by(conm, year) |> 
    summarise(avg_profit_margin = mean(`Profit Margin`, na.rm = TRUE), .groups = "drop") |>
    mutate(avg_profit_margin = round(avg_profit_margin * 100, 2),
           avg_profit_margin_label = percent(avg_profit_margin, accuracy = .01),
           year = as.factor(year),
           conm = as.factor(conm)) |> 
    select(avg_profit_margin, year, conm) |> 
    
    
    hchart("heatmap", 
           hcaes(x = year, y = conm, value = avg_profit_margin, name = conm, custom.year = year), 
    ) |>
    hc_add_theme(hc_theme_538()) |> 
    hc_colorAxis(stops = color_stops(colors = rev(c(
        "#0d47a1",
        "#1976d2",
        "#42a5f5", 
        "#ffb74d", 
        "#ff8a65",
        "#ff5722")))) |> 
    hc_xAxis(title = list(text = "Year")) |> 
    hc_yAxis(title = list(text = "")) |> 
    hc_title(text = '</span></strong></span> <span>&#128176;</span><span style="font-family: helvetica, inter, sans-serif;"><strong><span style="color: #black;"> Average Annual Profit Margin of Tech Giants',
             useHTML = TRUE) |> 
    hc_tooltip(
        xDateFormat = "%Y",
        headerFormat = "<b><span style='color:{series.color}; font-size:14px;'>{series.color}</span></b><br>",
        
        pointFormat = "<b>{point.name}</b> : {point.value:.2f}%",
        useHTML = TRUE)


matrix
    
```



---

## Q3 (3 points)

Create a line chart showing market capitalization over time for the top 5 companies by current market cap, with custom tooltips.


1. Find the top 5 companies by 2024 market cap and filter data for these companies
2. Filter data to years 2015-2024 and remove missing market cap values
3. Convert market cap to trillions for easier reading
4. Apply `hc_theme_economist()` theme
5. Create custom tooltip showing:
   - Company name in bold
   - Date formatted as "Month Year"
   - Market cap with "T" suffix and any relevant emoji you want to use. I used the company emoji 🏢



```{r}
#| label: Q3
#| fig-width: 12
#| fig-height: 8

# Find top 5 companies by latest market cap
q3 <- data2.0 |> 
    filter(year >= 2015, year <= 2024, conm %in% c(
        "ALPHABET INC", "AMAZON.COM INC", 
        "APPLE INC", "MICROSOFT CORP", 
        "NVIDIA CORP")) |>
    mutate(
        date = as.Date(datadate),
        mkt_cap = .000001 * (prccq * cshoq)) |> 
    hchart(
        "spline", 
        hcaes(x = date, y = mkt_cap, group = conm, value = conm),
        marker = list(enabled = TRUE)
    ) |> 
    hc_xAxis(
        type = "datetime",
        title = list(text = "Date")) |> 
    hc_yAxis(
        title = list(text = "Market Capitalization (Trillions $)")) |> 
    hc_add_theme(hc_theme_economist()) |> 
    hc_title(
        text = '<span>&#128200;</span> <strong style="color:black;">Market Cap Evolution of Tech Giants</strong>',
        useHTML = TRUE) |> 
    hc_tooltip(
        useHTML = TRUE,
        xDateFormat = "%A, %B %e, %Y",
        headerFormat = "<span style='font-size:12px;'>{point.key}</span><br>",
        pointFormat = "
      <span style='color:{series.color};font-weight:bold;'>
        {series.name}
      </span><br>
      Market Cap: <b>${point.y:.2f} T</b> 🏢")

q3
 
```

---

## Q4 (3 points)

Create a bubble chart showing the relationship between profit margin, asset turnover, and revenue (bubble size) for 2023 data.

1. Filter data to year 2023 and Q4 only (latest complete data)
2. Remove rows with missing profit margin, asset turnover, or revenue values
3. Filter out extreme outliers (profit margin between -0.5 and 1, asset turnover between 0 and 5)
4. Use `hchart()` with type "bubble"
5. Map x to asset turnover, y to profit margin, size to revenue (saleq), and group by company
6. Apply `hc_theme_flat()` theme
7. Set axis titles with HTML formatting and relevant emojis of you choice
8. Format tooltip to show company name, exact values, and revenue in millions

```{r}
#| label: Q4
#| fig-width: 11
#| fig-height: 8

q4 = data2.0 |> 
    filter(
        year == 2023,
        fqtr == 4, 
        `Profit Margin` >= -0.5, 
        `Profit Margin` <= 1,
        `Asset Turnover` >= 0,
        `Asset Turnover` <= 5) |> 
    mutate(`Profit Margin` = 100 * (`Profit Margin`)
    ) |> 
    hchart("bubble", hcaes(x = `Asset Turnover`, y = `Profit Margin`, group = conm, size = `Profit Margin`, name = conm, value = saleq)) |> 
    hc_add_theme(hc_theme_flat()) |> 
    hc_xAxis(
      title = list(text = "<span>💹</span> Asset Turnover (Revnue/Assets))")) |> 
    hc_yAxis(
    title = list(text = "<span>⚡</span> Profit Margin (Operating Income / Revenue)"),
    labels = list(format = "{value}%")) |>
    hc_add_theme(hc_theme_flat()) |> 
    hc_title(text = '</span></strong></span> <span>&#127919;</span><span style="font-family:Aptos Light;"><span style="color: #black;"> 2023 Q4: Efficiency vs Profitability (Bubble size = Revenue)',
             useHTML = TRUE) |> 
    hc_tooltip(
        headerFormat = "<span style='color:{series.color};font-weight:bold;'>
       {series.name}</span><br>",
        pointFormat = 
            "Asset Turnover: <b>{point.x:.2f} </b><br>
      Profit Margin: <b>{point.y:.1f}% </b><br>
      Revenue: <b>${point.value:.0f}M </b>",
        useHTML = TRUE
    )
q4



```

---

## Q5 (3 points)

Create a column chart showing year-over-year revenue growth for each company in 2023, with conditional coloring for positive/negative growth.


1. Calculate year-over-year revenue growth by comparing 2023 to 2022 annual revenues
2. Group by company and year, then summarize total annual revenue
3. Calculate the percentage growth from 2022 to 2023
4. Apply conditional coloring: green for positive growth, red for negative growth
5. Use `hc_theme_google()` theme
6. Customize the data labels to show percentage values on top of each column using `hc_plotOptions`


```{r}
#| label: Q5
#| fig-width: 11
#| fig-height: 8

q5 <- data2.0 |>
  filter(year %in% c(2022, 2023)) |>
  group_by(conm, year) |> 
  summarise(
    annual_revenue = sum(saleq, na.rm = TRUE), .groups = "drop") |> 
  tidyr::pivot_wider(
    names_from = year, values_from = annual_revenue, names_prefix = "rev_") |> 
  mutate(
    revenue_growth = (rev_2023 - rev_2022) / rev_2022 * 100,        
    color = ifelse(revenue_growth >= 0, "#4CAF50", "#f44336"),
    label = paste0("<b>", round(revenue_growth,1), "%</b>")) |> 
  hchart(
    "column", hcaes(
      x = conm, 
      y = revenue_growth, 
      color = color, 
      dataLabels = label)) |> 
  hc_plotOptions(
    column = list(
      dataLabels = list(
        enabled = TRUE, 
        useHTML = TRUE, 
        format = "<b>{point.y:.1f}%</b>")))|> 
  hc_add_theme(hc_theme_google()) |> 
  hc_xAxis(
    title = list(text = "")) |> 
    hc_yAxis(
    title = list(text = "Revenue Growth (%)"),
    labels = list(format = "{value}%"),
    tickInterval = 20) |> 
  hc_title(
    text = '<span style="font-family:Poppins, Helvetica, sans-serif;">2023 Revenue Growth: Winners and Losers</span>',
    useHTML = TRUE) |> 
  hc_subtitle(text = 'NVIDIA is killing it') |> 
  hc_tooltip(
    headerFormat = "<span style='font-size:12px;'>{point.key}</span><br>",
    pointFormat = "Revenue Growth: <b>{point.y:.1f}%</b>")

q5


```

---

