---
title: "Highcharts-Highcharter"
author: "Ashwin Malshe"
date: "21-October-2025"
format:
  html: 
    embed-resources: true
---

```{r}
#| label = "setup",
#| include = FALSE

pacman::p_load(highcharter, tidyverse)

```

JS visualizations add exponentially more functionality to static visualizations. As a very simple example, check out this visualization I made using highcharter:

https://dataviz.school/posts/2021-09-27-kiyosaki/

I will talk more about this in the class.


For the first exercise, we will use `citytemp` data from `highcharter` package.

```{r}
data(citytemp)
```


In a classic JS style, we treat each line on a plot below a series to be plotted separate of the X axis. We will again build it layer by layer similar to ggplot2. 

We first add the X axis and then lay down each line separately.

```{r}
hc = highchart() |> 
  hc_xAxis(categories = citytemp$month) |> 
  hc_add_series(name = "Berlin", data = citytemp$berlin) |>
  hc_add_series(name = "London", data = citytemp$london) |> 
  hc_add_series(name = "NYC", data = citytemp$new_york) |> 
  hc_add_series(name = "Tokyo", data = citytemp$tokyo)
hc

```

You can also follow a more traditional ggplot2-like structure by pivoting the data. Below, I use a custom function `hchart()` which is modeled on `ggplot()`. As you can see, there is a `hcaes()` similar to the `aes()` function in ggplot2.

However, you do not have to specify a separate `geom_*()` function here. Instead, you can pass the type of visualization you want to make (i.e., line graph in this case) as the second argument to `hchart()`. Note that the first argument to `hchart()` is an R object, which in our case is a dataframe called `citytemp`. 

```{r}
citytemp |> 
  pivot_longer(cols = where(is.numeric), names_to = "city", values_to = "temperature") |> 
  hchart(type = "line", 
         hcaes(x = month, y = temperature, group = city),  
         name = c("Berlin", "London", "NYC", "Tokyo"))
```

## Making alterations to the first plot we made above

I have liberally used highcharts documentation to make this plot. Notice how the title is matched with a list while other arguments have scalar parameters. This is because `htmlwidgets` converts an object initializer in JavaScript to an R list.  

"An object initializer is a comma-delimited list of zero or more pairs of property names and associated values of an object, enclosed in curly braces ({})." - https://developer.mozilla.org/en-US/docs/Web/JavaScript/Reference/Operators/Object_initializer



```{r}
hc |> 
  hc_yAxis(title = list(text = "Temperature on y Axis",
                        textAlign = "center",
                        style = list(color = "#776abc",
                                     fontSize = "1.8em",
                                     fontFamily = "Inter")),  # Replace with a font that's installed on your computer
           opposite = TRUE,
           minorTickInterval = "auto",
           minorGridLineDashStyle = "LongDashDotDot",
           showFirstLabel = T,
           showLastLabel = T) |> 
    hc_xAxis(title = list(text = "Month on x Axis",
                          style = list(color = "#776abc",
                                     fontSize = "1.8em",
                                     fontFamily = "American Typewriter")), # Replace with a font that's installed on your computer
           opposite = TRUE)
```
## Examples are from the author of the package

Most of these examples are from this website:

https://rpubs.com/jbkunst/230276

Highcharter has several themes. You can set the theme upfront as follows:

```{r}
#| eval = FALSE
options(highcharter.theme = hc_theme_smpl())
```

Above I am setting a simple theme. More themes are available here:
http://jkunst.com/highcharts-themes-collection/


Create a sample data frame

```{r}
df = data.frame(xval = 1:10) |> 
  mutate(
    yval = 10 + xval + 10 * sin(xval),
    yval = round(yval, 1),
    zval = (xval*yval) - median(xval*yval),
    er = 10 * abs(rnorm(length(xval))) + 2,
    er = round(er, 1),
    l = yval - er,
    h = yval + er,
    col = yval,
    nm = paste("point", xval)
  )

head(df)
```

```{r}

hchart(df, "column", hcaes(x = xval, y = zval, color = col, name = nm)) |> 
  hc_add_theme(hc_theme_monokai())
```
```{r}
hchart(df, "funnel", hcaes(y = yval, name = nm, color = col))
```
```{r}
hchart(df, "pie", hcaes(y = yval, name = nm, color = col))
```
```{r}
hchart(df, "pyramid", hcaes(y = yval, name = nm, color = col))
```

