---
title: "Themes"
format: html
---

```{r}
pacman::p_load(tidyverse, ggtext, hrbrthemes, ggthemes, showtext, ggedit)
theme_set(theme_minimal())
```


```{r}
dt = read_csv("company_financials_csv.zip", show_col_types = FALSE)
```


```{r}
base = dt |> 
  filter(!is.na(saleq) & !is.na(oiadpq)) |> 
  ggplot(aes(x = saleq, y = oiadpq)) +
  geom_point(shape = 21, color = "#bd2e38", fill = "#f82267", alpha = 0.6, size = 3) +
  labs(x = "Sales", y = "Operating Income") +
  scale_x_continuous(labels = scales::dollar_format()) +
  scale_y_continuous(labels = scales::dollar_format())
```

```{r}
base
```

A couple of themes from `ggthemes` package

```{r}
base + theme_economist()
```

```{r}
base + theme_fivethirtyeight()
```


A couple of themes from the `hrbrthemes` package. This package requires installing fonts to the system, which can be a bit cumbersome. A solution to this problem is to use `showtext`

We will use "Roboto Condensed" and "Inter" fonts.

```{r}
font_add_google(name = "Roboto Condensed", family = "rc")
font_add_google(name = "Inter", family = "inter")
```


```{r}
showtext_auto()
base + theme_ipsum_rc(base_family = "rc")
showtext_end()
```

Using Inter

```{r}
showtext_auto()
base +theme_ipsum_rc(base_family = "inter")
showtext_end()
```

Changing the Axes titles positioning, size, etc.

```{r}
showtext_auto()
base +
  theme_ipsum_rc(base_family = "inter",
                      axis_title_just = "cc",
                      axis_title_size = 16)
showtext_end()
```

Changing the font in the theme will only affect the text outside the panel. For example, notice the font in the annotation label.

```{r}

showtext_auto()
base +
  annotate(geom = "label", 
           x = 150000, y = 30000, 
           label = "There are two potential\nregression lines",
           color = "gray30") +
  theme_ipsum_rc(base_family = "rc")

showtext_end()
```

```{r}
showtext_auto()
base +
  annotate(geom = "label", 
           x = 150000, y = 30000, 
           label = "There are two potential\nregression lines",
           family = "rc", color = "gray30") +
  theme_ipsum_rc(base_family = "rc")
showtext_end()
```

A detour. Sometimes we want to remove a `geom` from an already created plot and replace it with another `geom`. ggedit package allows you to do this. For example, below I am remove the existing `geom_point` and replacing it with a new layer that adds `fill` aesthetic to differentiate points belonging to Amazon.com.

```{r}
showtext_auto()

base |> 
  ggedit::remove_geom("point") +
  geom_point(aes(fill = ifelse(tic == "AMZN", "1", "0")), 
             shape = 21, alpha = 0.6, size = 3, show.legend = FALSE) +
  annotate(geom = "label",
           x = 150000, y = 20000,
           label = "Amazon's profit margins\nare very low",
           family = "rc", color = "gray30") +
  theme_ipsum_rc(base_family = "rc")
showtext_end()
```